---
title: "Assignment 1: Research Questions"
author: "Alicia Seeley, JT Keller, Alec Martinez"
date: "10/20/2021"
output: 
  html_document:
    theme: readable
    toc: true
    toc_depth: 4
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Research Question to Answer with Regression

Linear: How, if at all, is household income, property tax, and age related to neighborhood social ties?

Logistic: How, if at all, is housing tenure related to neighborhood social ties. 

# Prior Research

Linear model: We have identified a study that 

We have identified two studies that show a positive relationship between housing tenure and different measures of social engagement. Levitten-Reid and Matthew (2018), using a sample from Canada, find "a positive association between home ownership and forms of bonding social capital". Similarly, Werner and Klingborg (2010), in a study conducted of Stockholm, Sweden, found results indicating that "residents forming tenant-owners associations have more trust in neighbors in the neighborhood."


# Sample Population and Variables

Our sample population includes adults in the US who participated in the IPUMS community civic engagement survey in November 2013.

## Categorial Variables

1. Housing Tenure

2. Participated in a service or civic organization

3. Trust in People in Neighborhood

4. Frequency of talking to neighbors

5. Frequency of doing favors for neighbors and vice versa

## Continuous Variables

1. Household Income

2. Age

3. Property Taxes


# Data Import

## Load Libraries

```{r, message=FALSE}
library(tidyverse)
library(tidycensus)
library(readxl)
library(knitr)
library(ipumsr)
library(survey)
library(srvyr)
library(ggplot2)
library(gridExtra)
```


## Read in .dat file from IPUMS

```{r, message=FALSE, warning=FALSE}
ddi <- read_ipums_ddi("cps_00004.xml")
data <- read_ipums_micro(ddi)
```

# Clean Data

```{r}
data <- data %>% drop_na(CENEIGHCONF)
```

```{r}
data = subset(data, select = -c(ASECFLAG,ASECWTH,GQ,OWNERSHP,HHINCOME,PROPTAX,PUBHOUS,UNITSSTR,ASECWT))
```


```{r}
kable(head(data))
```

# Descriptive Statistics

## Continuous variables

The continuous variables in our study are 1) household income, 2) age, 3) property tax.

### Calculations

#### t tests (including confidence interval)

```{r}
#income_t_test <- t.test(data$HHINCOME)
#age_t_test <- t.test(data$AGE)
#tax_t_test <- t.test(data$PROPTAX)
famincome_t_test <- t.test(data$FAMINC)
age_t_test <- t.test (data$AGE)

```


#### quantiles (including median and interquartile range)

```{r}
income_quartiles <- quantile(data$FAMINC, na.rm = TRUE)
age_quartiles <- quantile(data$AGE, na.rm = TRUE)

```


#### standard deviation

```{r}
income_sd <- sd(data$FAMINC, na.rm = TRUE)
age_sd <- sd(data$AGE, na.rm = TRUE)

```

#### histograms

```{r}
income_hist <- ggplot(data) +
  geom_histogram(aes(x = FAMINC)) 

#+
 # scale_x_continuous(trans = "log")

age_hist <- ggplot(data) +
  geom_histogram(aes(x = AGE))


```


### Displaying Data

```{r}
cont_summary <- tibble(
  Variable = c("Household income", 
               "Age"),
  `Sample mean` = c(famincome_t_test$estimate,
                    age_t_test$estimate),
  `Population mean (95% confidence) - low` = 
    c(famincome_t_test$conf.int[1],
      age_t_test$conf.int[1]),
  `Population mean (95% confidence) - high` =
    c(famincome_t_test$conf.int[2],
      age_t_test$conf.int[2]),
  Median = c(income_quartiles[3],
             age_quartiles[3]),
  `Interquartile range` = c(income_quartiles[4] - income_quartiles[2], age_quartiles[4] - age_quartiles[2]),
  `Standard deviation` = c(income_sd,
                          age_sd))

kable(cont_summary, digits = 0)
```



```{r}
pretty_income_hist <- income_hist +
  theme_bw() +
  scale_x_continuous(name = "Family Median income",
                     trans = "log") +
  scale_y_continuous(name = "Number of people") +
  theme(axis.text.x = element_text(angle = 90))

pretty_age_hist <- age_hist +
  theme_bw() + 
  scale_x_continuous(name = "Age") +
  scale_y_continuous(name = "Number of people") +
  theme(axis.text.x = element_text(angle = 90))

grid.arrange(pretty_income_hist, pretty_age_hist, ncol = 2)
```

```{r}
data <- data %>% 
  mutate(HHTENURE1 = case_when(`HHTENURE` == 0 ~ "NIU",
                          `HHTENURE` == 1 ~ "Owned or being bought",
                          `HHTENURE` == 2 ~ "Rented for cash",
                          `HHTENURE` == 3 ~ "Occupied without payment or cash rent",
                          `HHTENURE` == 6 ~ "Refused",
                          `HHTENURE` == 7 ~ "Don't know",
                          TRUE ~ "unknown"))
```

# STOP HERE

```{r}
pct_tenure_owned <- t.test(data$HHTENURE1 == "Owned or being bought")
pct_tenure_rented <- t.test(data$HHTENURE1 == "Rented for cash")
pct_tenure_freeload <-t.test(data$HHTENURE1 == "Occupied without payment or cash rent")
pct_tenure_other <-t.test(data$HHTENURE1 == c("NIU", "Refused", "Don't Know"))
```


```{r}
data <- data %>% 
  mutate(CENEIGHCONF1 = case_when(`CENEIGHCONF` == 0 ~ "NIU",
                          `CENEIGHCONF` == 1 ~ "Trust none of the People",
                          `CENEIGHCONF` == 2 ~ "Trust some of the People",
                          `CENEIGHCONF` == 3 ~ "Trust most of the People",
                          `CENEIGHCONF` == 6 ~ "Trust all of the People",
                          `CENEIGHCONF` == 96 ~ "Refused",
                          `CENEIGHCONF` == 97 ~ "Don't know",
                          `CENEIGHCONF` == 98 ~ "No Response",
                          `CENEIGHCONF` == 99 ~ "NIU",
                          TRUE ~ "unknown"))
```



```{r}
pct_trust_none <- t.test(data$CENEIGHCONF1 == "Trust none of the People")
pct_trust_some <- t.test(data$CENEIGHCONF1 == "Trust some of the People")
pct_trust_most <- t.test(data$CENEIGHCONF1 == "Trust most of the People")
pct_trust_all <- t.test(data$CENEIGHCONF1 == "Trust all of the People")
pct_trust_other <- t.test(data$CENEIGHCONF1 == c("Refused", "Don't know", "No Response", "NIU", "unknown"))
```


```{r}
data <- data %>% 
  mutate(CESOCIALIZE1 = case_when(`CESOCIALIZE` == 0 ~ "NIU",
                          `CESOCIALIZE` == 1 ~ "Not at all",
                          `CESOCIALIZE` == 2 ~ "Less than once a month",
                          `CESOCIALIZE` == 3 ~ "Once a month",
                          `CESOCIALIZE` == 4 ~ "A few times a month",
                          `CESOCIALIZE` == 5 ~ "A few times a week",
                          `CESOCIALIZE` == 6 ~ "Basically every day",
                          `CESOCIALIZE` == 96 ~ "Refused",
                          `CESOCIALIZE` == 97 ~ "Don't Know",
                          `CESOCIALIZE` == 98 ~ "No Response",
                          `CESOCIALIZE` == 99 ~ "NIU",
                          TRUE ~ "unknown"))
```


```{r}
pct_social_not <- t.test(data$CESOCIALIZE1 == "Not at all")
pct_social_less <- t.test(data$CESOCIALIZE1 == "Less than once a month")
pct_social_once <- t.test(data$CESOCIALIZE1 == "Once a month")
pct_social_few <- t.test(data$CESOCIALIZE1 == "A few times a month")
pct_social_week <- t.test(data$CESOCIALIZE1 == "A few times a week")
pct_social_day <- t.test(data$CESOCIALIZE1 == "Basically every day")
pct_social_other <- t.test(data$CESOCIALIZE1 == c("Refused", "Don't Know", "No Response", "NIU"))

```



```{r}
summary_tenure <-tibble(`Household Tenure` = 
                              c("Owned or being bought",
                                "Rented for cash",
                                "Occupied without payment or cash rent",
                                "NIU, Refused, Don't know"),
                            `Sample proportion` = 
                              c(pct_tenure_owned$estimate * 100,
                                pct_tenure_rented$estimate * 100,
                                pct_tenure_freeload$estimate * 100,
                                pct_tenure_other$estimate * 100),
                            `95-percent confidence - low` = 
                              c(pct_tenure_owned$conf.int[1] * 100,
                                pct_tenure_rented$conf.int[1] * 100,
                                pct_tenure_freeload$conf.int[1] * 100,
                                pct_tenure_other$conf.int[1] * 100),
                            `95-percent confidence - high` = 
                              c(pct_tenure_owned$conf.int[2] * 100,
                                pct_tenure_rented$conf.int[2] * 100,
                                pct_tenure_freeload$conf.int[2] * 100,
                                pct_tenure_other$conf.int[2] * 100))

kable(summary_tenure, digits = 0)
```

```{r}
summary_trust <-tibble(`Neighborhood Trust` = 
                              c("Trust none of the people",
                                "Trust some of the people",
                                "Trust most of the people",
                                "Trust all of the people",
                                "Refused, Don't Know, No Response, NIU"),
                            `Sample proportion` = 
                              c(pct_trust_none$estimate * 100,
                                pct_trust_some$estimate * 100,
                                pct_trust_most$estimate * 100,
                                pct_trust_all$estimate * 100,
                                pct_trust_other$estimate * 100),
                            `95-percent confidence - low` = 
                              c(pct_trust_none$conf.int[1] * 100,
                                pct_trust_some$conf.int[1] * 100,
                                pct_trust_most$conf.int[1] * 100,
                                pct_trust_all$conf.int[1] * 100,
                                pct_trust_other$conf.int[1] * 100),
                            `95-percent confidence - high` = 
                              c(pct_trust_none$conf.int[2] * 100,
                                pct_trust_some$conf.int[2] * 100,
                                pct_trust_most$conf.int[2] * 100,
                                pct_trust_all$conf.int[2] * 100,
                                pct_trust_other$conf.int[2] * 100))

kable(summary_trust, digits = 0)
```
```{r}
summary_social <-tibble(`Social Neighbors` = 
                              c("Not at all",
                                "Less than once a month",
                                "Once a month",
                                "A few times a month",
                                "A few times a week",
                                "Basically every day",
                                "Refused, Don't Know, No Response, NIU"),
                            `Sample proportion` = 
                              c(pct_social_not$estimate * 100,
                                pct_social_less$estimate * 100,
                                pct_social_once$estimate * 100,
                                pct_social_few$estimate * 100,
                                pct_social_week$estimate * 100,
                                pct_social_day$estimate * 100,
                                pct_social_other$estimate * 100),
                            `95-percent confidence - low` = 
                              c(pct_social_not$conf.int[1] * 100,
                                pct_social_less$conf.int[1] * 100,
                                pct_social_once$conf.int[1] * 100,
                                pct_social_few$conf.int[1] * 100,
                                pct_social_week$conf.int[1] * 100,
                                pct_social_day$conf.int[1] * 100,
                                pct_social_other$conf.int[1] * 100),
                            `95-percent confidence - high` = 
                              c(pct_social_not$conf.int[2] * 100,
                                pct_social_less$conf.int[2] * 100,
                                pct_social_once$conf.int[2] * 100,
                                pct_social_few$conf.int[2] * 100,
                                pct_social_week$conf.int[2] * 100,
                                pct_social_day$conf.int[2] * 100,
                                pct_social_other$conf.int[2] * 100))

kable(summary_social, digits = 0)
```



# References

Werner, Inga Britt, and Kerstin Klingborg. "Studying Social Capital In Housing Neighborhoods-Does Tenure Matter?." (2010).

Leviten-Reid, Catherine, and Rebecca A. Matthew. "Housing tenure and neighbourhood social capital." Housing, Theory and Society 35, no. 3 (2018): 300-328.

Sarah Flood, Miriam King, Renae Rodgers, Steven Ruggles, J. Robert Warren and Michael Westberry. Integrated Public Use Microdata Series, Current Population Survey: Version 9.0 [dataset]. Minneapolis, MN: IPUMS, 2021. https://doi.org/10.18128/D030.V9.0
