---
title: "Assignment 1: Research Questions"
author: "Alicia Seeley, JT Keller, Alec Martinez"
date: "10/20/2021"
output: 
  html_document:
    theme: readable
    toc: true
    toc_depth: 4
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Research Question to Answer with Regression

Linear: How, if at all, is household income, property tax, and age related to neighborhood social ties?

Logistic: How, if at all, is housing tenure related to neighborhood social ties. 

# Prior Research

Linear model: We have identified a study that 

We have identified two studies that show a positive relationship between housing tenure and different measures of social engagement. Levitten-Reid and Matthew (2018), using a sample from Canada, find "a positive association between home ownership and forms of bonding social capital". Similarly, Werner and Klingborg (2010), in a study conducted of Stockholm, Sweden, found results indicating that "residents forming tenant-owners associations have more trust in neighbors in the neighborhood."


# Sample Population and Variables

Our sample population includes adults in the US who participated in the IPUMS community civic engagement survey in November 2013.

## Categorial Variables

### 1. Housing Tenure

### 2. Participated in a service or civic organization

### 3. Trust in People in Neighborhood

### 4. Frequency of talking to neighbors

### 5. Frequency of doing favors for neighbors and vice versa

## Continuous Variables

### 1. Household Income

### 2. Age

### 3. Property Taxes


# Data Import

## Load Libraries

```{r, message=FALSE}
library(tidyverse)
library(tidycensus)
library(readxl)
library(knitr)
library(ipumsr)
library(survey)
library(srvyr)
library(ggplot2)
library(gridExtra)
```


## Read in .dat file from IPUMS

```{r, message=FALSE, warning=FALSE}
ddi <- read_ipums_ddi("cps_00003.xml")
data <- read_ipums_micro(ddi)
```


```{r}
svy_data <- data %>%
  as_survey_design(ids = NULL,
                   strata = NULL,
                   weights = HWTFINL)
```



```{r}
kable(head(data))
```

# Descriptive Statistics

## Continuous variables

The continuous variables in our study are 1) household income, 2) age, 3) property tax.

### Calculations

#### t tests (including confidence interval)

```{r}
income_t_test <- t.test(data$HHINCOME)
age_t_test <- t.test(data$AGE)
tax_t_test <- t.test(data$PROPTAX)

```



#### quantiles (including median and interquartile range)

```{r}
income_quartiles <- quantile(data$HHINCOME, na.rm = TRUE)
age_quartiles <- quantile(data$AGE, na.rm = TRUE)
tax_quartiles <- quantile(data$PROPTAX, na.rm = TRUE)

```


#### standard deviation

```{r}
income_sd <- sd(data$HHINCOME, na.rm = TRUE)
age_sd <- sd(data$AGE, na.rm = TRUE)
tax_sd <- sd(data$PROPTAX, na.rm = TRUE)

```

#### histograms

```{r}
income_hist <- ggplot(data) +
  geom_histogram(aes(x = HHINCOME)) +
  scale_x_continuous(trans = "log")

age_hist <- ggplot(data) +
  geom_histogram(aes(x = AGE))

tax_hist <- ggplot(data) +
  geom_histogram(aes(x = PROPTAX)) +
  scale_x_continuous(trans = "log")
```



### Displaying Data

```{r}
cont_summary <- tibble(
  Variable = c("Household income", 
               "Age", 
               "Property tax"),
  `Sample mean` = c(income_t_test$estimate,
                    age_t_test$estimate,
                    tax_t_test$estimate),
  `Population mean (95% confidence) - low` = 
    c(income_t_test$conf.int[1],
      age_t_test$conf.int[1],
      tax_t_test$conf.int[1]),
  `Population mean (95% confidence) - high` =
    c(income_t_test$conf.int[2],
      age_t_test$conf.int[2],
      tax_t_test$conf.int[2]),
  Median = c(income_quartiles[3],
             age_quartiles[3],
             tax_quartiles[3]),
  `Interquartile range` = c(income_quartiles[4] - income_quartiles[2],
                            age_quartiles[4] - age_quartiles[2],
                            (tax_quartiles[4] - tax_quartiles[2])),
  `Standard deviation` = c(income_sd,
                          age_sd,
                          tax_sd))

kable(cont_summary, digits = 0)
```
```{r}
pretty_income_hist <- income_hist +
  theme_bw() +
  scale_x_continuous(name = "Median income",
                     trans = "log") +
  scale_y_continuous(name = "Number of people") +
  theme(axis.text.x = element_text(angle = 90))

pretty_age_hist <- age_hist +
  theme_bw() + 
  scale_x_continuous(name = "Age") +
  scale_y_continuous(name = "Number of people") +
  theme(axis.text.x = element_text(angle = 90))

pretty_tax_hist = tax_hist +
  theme_bw() +
  scale_x_continuous(name = "Property Taxes",
                     trans = "log") +
  scale_y_continuous(name = "Number of people") +
  theme(axis.text.x = element_text(angle = 90)) 

grid.arrange(pretty_income_hist, pretty_age_hist, pretty_tax_hist,
             ncol = 3)
```
```{r}
data <- data %>% 
  mutate(HHTENURE = case_when(`HHTENURE` == 0 ~ "NIU",
                          `HHTENURE` == 1 ~ "Owned or being bought",
                          `HHTENURE` == 2 ~ "Rented for cash",
                          `HHTENURE` == 3 ~ "Occupied without payment or cash rent",
                          `HHTENURE` == 6 ~ "Refused",
                          `HHTENURE` == 7 ~ "Don't know",
                          TRUE ~ "unknown"))

data <- data %>% 
  mutate(CENEIGHCONF = case_when(`CENEIGHCONF` == 0 ~ "NIU",
                          `CENEIGHCONF` == 1 ~ "Trust none of the People",
                          `CENEIGHCONF` == 2 ~ "Trust some of the People",
                          `CENEIGHCONF` == 3 ~ "Trust most of the People",
                          `CENEIGHCONF` == 6 ~ "Trust all of the People",
                          `CENEIGHCONF` == 96 ~ "Refused",
                          `CENEIGHCONF` == 97 ~ "Don't know",
                          `CENEIGHCONF` == 98 ~ "No Response",
                          `CENEIGHCONF` == 99 ~ "NIU",
                          TRUE ~ "unknown"))

data <- data %>% 
  mutate(CESOCIALIZE = case_when(`CESOCIALIZE` == 0 ~ "NIU",
                          `CESOCIALIZE` == 1 ~ "Not at all",
                          `CESOCIALIZE` == 2 ~ "Less than once a month",
                          `CESOCIALIZE` == 3 ~ "Once a month",
                          `CESOCIALIZE` == 4 ~ "A few times a month",
                          `CESOCIALIZE` == 5 ~ "A few times a week",
                          `CESOCIALIZE` == 6 ~ "Basically every day",
                          `CESOCIALIZE` == 96 ~ "Refused",
                          `CESOCIALIZE` == 97 ~ "Don't Know",
                          `CESOCIALIZE` == 98 ~ "No Response",
                          `CESOCIALIZE` == 99 ~ "NIU",
                          TRUE ~ "unknown"))
```


```{r}

```


# References

Werner, Inga Britt, and Kerstin Klingborg. "Studying Social Capital In Housing Neighborhoods-Does Tenure Matter?." (2010).

Leviten-Reid, Catherine, and Rebecca A. Matthew. "Housing tenure and neighbourhood social capital." Housing, Theory and Society 35, no. 3 (2018): 300-328.

Sarah Flood, Miriam King, Renae Rodgers, Steven Ruggles, J. Robert Warren and Michael Westberry. Integrated Public Use Microdata Series, Current Population Survey: Version 9.0 [dataset]. Minneapolis, MN: IPUMS, 2021. https://doi.org/10.18128/D030.V9.0
