---
title: "Assignment 1: Research Questions"
author: "Alicia Seeley, JT Keller, Alec Martinez"
date: "10/20/2021"
output: 
  html_document:
    theme: readable
    toc: true
    toc_depth: 4
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Research Question to Answer with Regression

Linear: How, if at all, is household income, property tax, and age related to neighborhood social ties?

Logistic: How, if at all, is housing tenure related to neighborhood social ties. 

# Prior Research

Linear model: We have identified a study that 

We have identified two studies that show a positive relationship between housing tenure and different measures of social engagement. Levitten-Reid and Matthew (2018), using a sample from Canada, find "a positive association between home ownership and forms of bonding social capital". Similarly, Werner and Klingborg (2010), in a study conducted of Stockholm, Sweden, found results indicating that "residents forming tenant-owners associations have more trust in neighbors in the neighborhood."


# Sample Population and Variables

Our sample population includes adults in the US who participated in the IPUMS community civic engagement survey in November 2013.

## Categorial Variables

1. Housing Tenure

2. Participated in a service or civic organization

3. Trust in People in Neighborhood

4. Frequency of talking to neighbors

5. Frequency of doing favors for neighbors and vice versa

## Continuous Variables

1. Household Income

2. Age

3. Property Taxes


# Data Import

## Load Libraries

```{r, message=FALSE}
library(tidyverse)
library(tidycensus)
library(readxl)
library(knitr)
library(ipumsr)
library(survey)
library(srvyr)
library(ggplot2)
library(gridExtra)
library(nnet)
```


## Read in .dat file from IPUMS

```{r, message=FALSE, warning=FALSE}
ddi <- read_ipums_ddi("cps_00004.xml")
data <- read_ipums_micro(ddi)
```

# Clean Data

```{r}
data <- data %>% drop_na(CENEIGHCONF)
```

```{r}
data <- data %>% filter(CENEIGHCONF != 99)
```

```{r}
data <- data %>% filter(CENEIGHCONF != 96)
```

```{r}
data <- data %>% filter(CENEIGHCONF != 97)
```

```{r}
data <- data %>% filter(CENEIGHCONF != 98)
```

```{r}
data = subset(data, select = -c(ASECFLAG,ASECWTH,GQ,OWNERSHP,HHINCOME,PROPTAX,PUBHOUS,UNITSSTR,ASECWT))
```


```{r}
kable(head(data))
```






```{r}
data <- data %>% 
  mutate(HHTENURE1 = case_when(`HHTENURE` == 0 ~ "NIU",
                          `HHTENURE` == 1 ~ "Owned or being bought",
                          `HHTENURE` == 2 ~ "Rented for cash",
                          `HHTENURE` == 3 ~ "Occupied without payment or cash rent",
                          `HHTENURE` == 6 ~ "Refused",
                          `HHTENURE` == 7 ~ "Don't know",
                          TRUE ~ "unknown"))
```

# STOP HERE

```{r}
pct_tenure_owned <- t.test(data$HHTENURE1 == "Owned or being bought")
pct_tenure_rented <- t.test(data$HHTENURE1 == "Rented for cash")
pct_tenure_freeload <-t.test(data$HHTENURE1 == "Occupied without payment or cash rent")
pct_tenure_other <-t.test(data$HHTENURE1 == c("NIU", "Refused", "Don't Know"))
```

#DESCRIPTIVE STATISTICS

##CONTINUOUS VARIABLES

Percent Renters
Trust

```{r}
pct_rent_mean <- all_data %>%
  srvyr::summarize(mean = survey_mean(pct_rent, vartype = c("ci", "var"), na.rm = TRUE))

trust_mean <- all_data %>%
  srvyr::summarize(mean = survey_mean(mean, vartype = c("ci", "var"), na.rm = TRUE))
```

```{r}
rent_quants <- all_data %>%
  srvyr::summarize(quants = survey_quantile(pct_rent, 
                                     quantiles =  c(0.25, 0.5, 0.75),
                                     vartype = NULL)) 

trust_quants <- all_data %>%
  srvyr::summarize(quants = survey_quantile(mean, 
                                     quantiles =  c(0.25, 0.5, 0.75),
                                     vartype = NULL)) 
```

```{r}
mean_summary <- rbind(pct_rent_mean, trust_mean) %>%
    mutate(variable = c("Percentage of Renters",
                      "Percentage that trust Neighbors"))

quant_summary <- rbind(rent_quants, trust_quants) %>%
    mutate(variable = c("Percentage of Renters",
                      "Percentage that trust Neighbors"))

summary <- left_join(mean_summary, quant_summary) %>%
  mutate(sd = mean_var^0.5) %>%
  mutate(IQR = quants_q75 - quants_q25) %>%
  rename(median = quants_q50) %>%
  select(variable, mean, mean_low, mean_upp, median, sd, IQR)

kable(summary, digits=2)
```

```{r}
renters_hist <- svyhist(~pct_rent, design = all_data,
                            freq = TRUE)
```

```{r}
trust_hist <- svyhist(~mean, design = all_data,
                            freq = TRUE)
```

```{r}
renters_hist_df <- tibble(mids = renters_hist$mids,
                              counts = renters_hist$counts)

ggplot(renters_hist_df) +
  geom_bar(aes(x = mids, y = counts),
           stat = "identity") +
  theme_bw() +
  scale_x_continuous(name = "Percentage of Renters") +
  scale_y_continuous(name = "Estimated number of people?")
```

```{r}
trust_hist_df <- tibble(mids = trust_hist$mids,
                              counts = trust_hist$counts)

ggplot(trust_hist_df) +
  geom_bar(aes(x = mids, y = counts),
           stat = "identity") +
  theme_bw() +
  scale_x_continuous(name = "Percentage of People who trust neighbors") +
  scale_y_continuous(name = "Estimated number of People")
```

##CATAGORICAL VARIABLE

Income

```{r}
inc_shares <- all_data %>%
  group_by(INCOME_BINS) %>%
  srvyr::summarize(share = survey_prop(vartype = "ci"))

kable(inc_shares, digits = 3)
```

# CORRELATIONS ON CONTINUOUS VARIABLES

```{r}
corrs <- svycor(~pct_rent + mean, design = all_data, sig.stats = TRUE)
```

```{r}
corrs$cors
```

```{r}
corrs$p.values
```

```{r}
corrs$std.err
```

```{r}
corrs_ci_low <- corrs$cors - 1.96*corrs$std.err

corrs_ci_upp <- corrs$cors + 1.96*corrs$std.err
```

```{r}
corrs_ci_low
```

```{r}
corrs_ci_upp
```

#REGRESSION

```{r}
tenure_trust_model <- svyglm(pct_rent ~ mean, design = all_data)

summary(tenure_model)
```

```{r}
income_trust_model <- svyglm(INCOME_BINS ~ mean, design = all_data)

summary(income_model)
```











```{r}
data <- data %>% 
  mutate(CENEIGHCONF1 = case_when(`CENEIGHCONF` == 0 ~ "NIU",
                          `CENEIGHCONF` == 1 ~ "Trust none of the People",
                          `CENEIGHCONF` == 2 ~ "Trust some of the People",
                          `CENEIGHCONF` == 3 ~ "Trust most of the People",
                          `CENEIGHCONF` == 6 ~ "Trust all of the People",
                          `CENEIGHCONF` == 96 ~ "Refused",
                          `CENEIGHCONF` == 97 ~ "Don't know",
                          `CENEIGHCONF` == 98 ~ "No Response",
                          `CENEIGHCONF` == 99 ~ "NIU",
                          TRUE ~ "unknown"))
```



```{r}
pct_trust_none <- t.test(data$CENEIGHCONF1 == "Trust none of the People")
pct_trust_some <- t.test(data$CENEIGHCONF1 == "Trust some of the People")
pct_trust_most <- t.test(data$CENEIGHCONF1 == "Trust most of the People")
pct_trust_all <- t.test(data$CENEIGHCONF1 == "Trust all of the People")
pct_trust_other <- t.test(data$CENEIGHCONF1 == c("Refused", "Don't know", "No Response", "NIU", "unknown"))
```


```{r}
data <- data %>% 
  mutate(CESOCIALIZE1 = case_when(`CESOCIALIZE` == 0 ~ "NIU",
                          `CESOCIALIZE` == 1 ~ "Not at all",
                          `CESOCIALIZE` == 2 ~ "Less than once a month",
                          `CESOCIALIZE` == 3 ~ "Once a month",
                          `CESOCIALIZE` == 4 ~ "A few times a month",
                          `CESOCIALIZE` == 5 ~ "A few times a week",
                          `CESOCIALIZE` == 6 ~ "Basically every day",
                          `CESOCIALIZE` == 96 ~ "Refused",
                          `CESOCIALIZE` == 97 ~ "Don't Know",
                          `CESOCIALIZE` == 98 ~ "No Response",
                          `CESOCIALIZE` == 99 ~ "NIU",
                          TRUE ~ "unknown"))
```


```{r}
pct_social_not <- t.test(data$CESOCIALIZE1 == "Not at all")
pct_social_less <- t.test(data$CESOCIALIZE1 == "Less than once a month")
pct_social_once <- t.test(data$CESOCIALIZE1 == "Once a month")
pct_social_few <- t.test(data$CESOCIALIZE1 == "A few times a month")
pct_social_week <- t.test(data$CESOCIALIZE1 == "A few times a week")
pct_social_day <- t.test(data$CESOCIALIZE1 == "Basically every day")
pct_social_other <- t.test(data$CESOCIALIZE1 == c("Refused", "Don't Know", "No Response", "NIU"))

```



```{r}
summary_tenure <-tibble(`Household Tenure` = 
                              c("Owned or being bought",
                                "Rented for cash",
                                "Occupied without payment or cash rent",
                                "NIU, Refused, Don't know"),
                            `Sample proportion` = 
                              c(pct_tenure_owned$estimate * 100,
                                pct_tenure_rented$estimate * 100,
                                pct_tenure_freeload$estimate * 100,
                                pct_tenure_other$estimate * 100),
                            `95-percent confidence - low` = 
                              c(pct_tenure_owned$conf.int[1] * 100,
                                pct_tenure_rented$conf.int[1] * 100,
                                pct_tenure_freeload$conf.int[1] * 100,
                                pct_tenure_other$conf.int[1] * 100),
                            `95-percent confidence - high` = 
                              c(pct_tenure_owned$conf.int[2] * 100,
                                pct_tenure_rented$conf.int[2] * 100,
                                pct_tenure_freeload$conf.int[2] * 100,
                                pct_tenure_other$conf.int[2] * 100))

kable(summary_tenure, digits = 0)
```

```{r}
summary_trust <-tibble(`Neighborhood Trust` = 
                              c("Trust none of the people",
                                "Trust some of the people",
                                "Trust most of the people",
                                "Trust all of the people",
                                "Refused, Don't Know, No Response, NIU"),
                            `Sample proportion` = 
                              c(pct_trust_none$estimate * 100,
                                pct_trust_some$estimate * 100,
                                pct_trust_most$estimate * 100,
                                pct_trust_all$estimate * 100,
                                pct_trust_other$estimate * 100),
                            `95-percent confidence - low` = 
                              c(pct_trust_none$conf.int[1] * 100,
                                pct_trust_some$conf.int[1] * 100,
                                pct_trust_most$conf.int[1] * 100,
                                pct_trust_all$conf.int[1] * 100,
                                pct_trust_other$conf.int[1] * 100),
                            `95-percent confidence - high` = 
                              c(pct_trust_none$conf.int[2] * 100,
                                pct_trust_some$conf.int[2] * 100,
                                pct_trust_most$conf.int[2] * 100,
                                pct_trust_all$conf.int[2] * 100,
                                pct_trust_other$conf.int[2] * 100))

kable(summary_trust, digits = 0)
```


```{r}
summary_social <-tibble(`Social Neighbors` = 
                              c("Not at all",
                                "Less than once a month",
                                "Once a month",
                                "A few times a month",
                                "A few times a week",
                                "Basically every day",
                                "Refused, Don't Know, No Response, NIU"),
                            `Sample proportion` = 
                              c(pct_social_not$estimate * 100,
                                pct_social_less$estimate * 100,
                                pct_social_once$estimate * 100,
                                pct_social_few$estimate * 100,
                                pct_social_week$estimate * 100,
                                pct_social_day$estimate * 100,
                                pct_social_other$estimate * 100),
                            `95-percent confidence - low` = 
                              c(pct_social_not$conf.int[1] * 100,
                                pct_social_less$conf.int[1] * 100,
                                pct_social_once$conf.int[1] * 100,
                                pct_social_few$conf.int[1] * 100,
                                pct_social_week$conf.int[1] * 100,
                                pct_social_day$conf.int[1] * 100,
                                pct_social_other$conf.int[1] * 100),
                            `95-percent confidence - high` = 
                              c(pct_social_not$conf.int[2] * 100,
                                pct_social_less$conf.int[2] * 100,
                                pct_social_once$conf.int[2] * 100,
                                pct_social_few$conf.int[2] * 100,
                                pct_social_week$conf.int[2] * 100,
                                pct_social_day$conf.int[2] * 100,
                                pct_social_other$conf.int[2] * 100))

kable(summary_social, digits = 0)
```
# Bivariate Analysis

```{r}
cor.test(~ CENEIGHCONF + HHTENURE, data = data)
```

```{r}
chisq.test(data$CENEIGHCONF, data$HHTENURE, correct = FALSE)
```
```{r}
chisq.test(data$CESOCIALIZE, data$HHTENURE, correct = FALSE)
```
```{r}
#data$CENEIGHCONF2 <- relevel(data$CENEIGHCONF1, ref = "Trust None of the People")
test <- multinom(CENEIGHCONF1 ~ HHTENURE1, data = data)
```


```{r}
summary(test)
e ```



# References

Werner, Inga Britt, and Kerstin Klingborg. "Studying Social Capital In Housing Neighborhoods-Does Tenure Matter?." (2010).

Leviten-Reid, Catherine, and Rebecca A. Matthew. "Housing tenure and neighbourhood social capital." Housing, Theory and Society 35, no. 3 (2018): 300-328.

Sarah Flood, Miriam King, Renae Rodgers, Steven Ruggles, J. Robert Warren and Michael Westberry. Integrated Public Use Microdata Series, Current Population Survey: Version 9.0 [dataset]. Minneapolis, MN: IPUMS, 2021. https://doi.org/10.18128/D030.V9.0
