---
title: "Assignment 1: Research Questions"
author: "Alicia Seeley, JT Keller, Alec Martinez"
date: "10/20/2021"
output: 
  html_document:
    theme: readable
    toc: true
    toc_depth: 4
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Research Question to Answer with Regression

How, if at all, is housing tenure related to neighborhood social ties. 

# Prior Research

We have identified two studies that show a positive relationship between housing tenure and different measures of social engagement. Levitten-Reid and Matthew (2018), using a sample from Canada, find "a positive association between home ownership and forms of bonding social capital". Similarly, Werner and Klingborg (2010), in a study conducted of Stockholm, Sweden, found results indicating that "residents forming tenant-owners associations have more trust in neighbors in the neighborhood."


# Sample Population and Variables

Our sample population includes adults in the US who participated in the IPUMS community civic engagement survey in November 2013.

## Continuous Variables (adapted from categorical variables)

1. Housing Tenure

2. Trust in People in Neighborhood



## Categorical Variables (adapted from continuous variables)

1. Household Income




# Data Import

## Load Libraries

```{r, message=FALSE}
library(tidyverse)
library(tidycensus)
library(readxl)
library(knitr)
library(ipumsr)
library(survey)
library(srvyr)
library(ggplot2)
library(gridExtra)
```


## Read in .dat file from IPUMS

```{r, message=FALSE, warning=FALSE}
ddi <- read_ipums_ddi("cps_00004.xml")
data <- read_ipums_micro(ddi)
```

# Clean Data

```{r}
data <- data %>% drop_na(CENEIGHCONF)
```


```{r}
data <- data %>% 
  mutate(CENEIGHCONF1 = case_when(`CENEIGHCONF` == 0 ~ "NIU",
                          `CENEIGHCONF` == 1 ~ "Trust none of the People",
                          `CENEIGHCONF` == 2 ~ "Trust some of the People",
                          `CENEIGHCONF` == 3 ~ "Trust most of the People",
                          `CENEIGHCONF` == 6 ~ "Trust all of the People",
                          `CENEIGHCONF` == 96 ~ "Refused",
                          `CENEIGHCONF` == 97 ~ "Don't know",
                          `CENEIGHCONF` == 98 ~ "No Response",
                          `CENEIGHCONF` == 99 ~ "NIU",
                          TRUE ~ "unknown"))
```


```{r}
data_subset <- data %>%
  filter(CENEIGHCONF1 != "NIU") %>%
  mutate(conf_2level = case_when(CENEIGHCONF1 == "Trust most of the People" ~ TRUE,
                                 CENEIGHCONF1 == "Trust all of the People" ~ TRUE,
                                 TRUE ~ FALSE))

svy_data <- data_subset %>%
  as_survey_design(weights = WTFINL)


county_level <- svy_data %>%
  group_by(COUNTY) %>%
  srvyr::summarize(mean = survey_mean(conf_2level),
                   sample = n()) 
```



Get census data

```{r}
acs_vars <- load_variables(2019, "acs5")
```


```{r}
county_data <- get_acs(geography = "county", variables = c("B25003_003", "B25003_001", "B06011_001"), output = 'wide') %>%
  mutate(pct_rent = B25003_003E / B25003_001E)

```
```{r}
county_data <- county_data %>% 
  rename("median_income" = "B06011_001E")

```

```{r}
income_quartile <- quantile(county_data$median_income, na.rm = TRUE)

income_quartile
```
```{r}
county_data$income_bins <- cut(county_data$median_income, breaks = c(0, 24000, 31000, Inf),
                               labels = c("bottom quartile", "middle quartile", "upper quartile"))
```


```{r}
county_level <- county_level %>% 
mutate(GEOID = case_when(str_length(as.character(COUNTY)) == 5 ~ 
                            as.character(COUNTY),
                          str_length(as.character(COUNTY)) == 4 ~
                            paste("0", COUNTY, sep=""),
                          TRUE ~ "not a county")) %>%
  filter(GEOID != "not a county")
```

### FIX 4-digit FIPS codes by putting the leading zeros back on them,
### the resulting FIPS code needs to be a string (as.character())
### Name the resulting FIPS code GEOID



```{r}
all_data <- left_join(county_level, county_data)

```

```{r}
kable(head(all_data))
```

## Descriptive Statistics

### Continuous Variables


```{r}
rent_ttest <- t.test(all_data$pct_rent)

trust_ttest <- t.test(all_data$mean)
```

```{r}
rent_quartiles <- quantile(all_data$pct_rent, na.rm = TRUE)
trust_quartiles <- quantile(all_data$mean, na.rm = TRUE)
```

```{r}
rent_sd <- sd(all_data$pct_rent, na.rm = TRUE)
trust_sd <- sd(all_data$mean, na.rm = TRUE)
```

```{r}
rent_hist <- ggplot(all_data) +
  geom_histogram(aes(x = pct_rent),
                 bins = 30)
```

```{r}
trust_hist <- ggplot(all_data) +
  geom_histogram(aes(x = mean),
                 bins = 30)
```

```{r}
summary <- tibble(
  Variable = c("Percentage of people who trust their neighbors (by county)", 
               "Percentage of renters (by county)"),
  `Sample mean` = c(trust_ttest$estimate,
                    rent_ttest$estimate),
  `Population mean (95% confidence) - low` = 
    c(trust_ttest$conf.int[1],
      rent_ttest$conf.int[1]),
  `Population mean (95% confidence) - high` =
    c(trust_ttest$conf.int[2],
      rent_ttest$conf.int[2]),
  Median = c(trust_quartiles[3],
             rent_quartiles[3]),
  `Interquartile range` = c(trust_quartiles[4] - trust_quartiles[2],
                            rent_quartiles[4] - rent_quartiles[2]),
  `Standard deviation` = c(trust_sd,
                          rent_sd))

kable(summary, digits = 2)
```

```{r}
pretty_trust_hist <- trust_hist +
  theme_bw() +
  scale_x_continuous(name = "percentage of ppl who trust neighbors") +
  scale_y_continuous(name = "Number of counties") +
  theme(axis.text.x = element_text(angle = 90))

pretty_rent_hist <- rent_hist +
  theme_bw() + 
  scale_x_continuous(name = "Percentage of renters") +
  scale_y_continuous(name = "Number of counties") +
  theme(axis.text.x = element_text(angle = 90))

grid.arrange(pretty_trust_hist, pretty_rent_hist,
             ncol = 2)
```

### Categorical variables

Still cleaning our categorical variables after starting over from scratch.

## Bivariate Analysis

### continuous variables

```{r}
cor.test(~ mean + pct_rent, data = all_data)

```

```{r}
trust_model <- lm(mean ~ pct_rent, data = all_data)

summary(trust_model)
```

### Categorical Variables

(Sorry that our categorical variable is still WIP)


# References

Werner, Inga Britt, and Kerstin Klingborg. "Studying Social Capital In Housing Neighborhoods-Does Tenure Matter?." (2010).

Leviten-Reid, Catherine, and Rebecca A. Matthew. "Housing tenure and neighbourhood social capital." Housing, Theory and Society 35, no. 3 (2018): 300-328.

Sarah Flood, Miriam King, Renae Rodgers, Steven Ruggles, J. Robert Warren and Michael Westberry. Integrated Public Use Microdata Series, Current Population Survey: Version 9.0 [dataset]. Minneapolis, MN: IPUMS, 2021. https://doi.org/10.18128/D030.V9.0
