---
title: "Assignment 1: Research Questions"
author: "Alicia Seeley, JT Keller, Alec Martinez"
date: "10/20/2021"
output: 
  html_document:
    theme: readable
    toc: true
    toc_depth: 4
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Research Question to Answer with Regression

Logistic: How, if at all, is housing tenure related to neighborhood social ties. 

# Prior Research

Linear model: We have identified a study that 

We have identified two studies that show a positive relationship between housing tenure and different measures of social engagement. Levitten-Reid and Matthew (2018), using a sample from Canada, find "a positive association between home ownership and forms of bonding social capital". Similarly, Werner and Klingborg (2010), in a study conducted of Stockholm, Sweden, found results indicating that "residents forming tenant-owners associations have more trust in neighbors in the neighborhood."


# Sample Population and Variables

Our sample population includes adults in the US who participated in the IPUMS community civic engagement survey in November 2013.

## Categorial Variables

1. Housing Tenure

2. Participated in a service or civic organization

3. Trust in People in Neighborhood

4. Frequency of talking to neighbors

5. Frequency of doing favors for neighbors and vice versa

## Continuous Variables

1. Household Income

2. Age

3. Property Taxes


# Data Import

## Load Libraries

```{r, message=FALSE}
library(tidyverse)
library(tidycensus)
library(readxl)
library(knitr)
library(ipumsr)
library(survey)
library(srvyr)
library(ggplot2)
library(gridExtra)
```


## Read in .dat file from IPUMS

```{r, message=FALSE, warning=FALSE}
ddi <- read_ipums_ddi("cps_00004.xml")
data <- read_ipums_micro(ddi)
```

# Clean Data

```{r}
data <- data %>% drop_na(CENEIGHCONF)
```


```{r}
data <- data %>% 
  mutate(CENEIGHCONF1 = case_when(`CENEIGHCONF` == 0 ~ "NIU",
                          `CENEIGHCONF` == 1 ~ "Trust none of the People",
                          `CENEIGHCONF` == 2 ~ "Trust some of the People",
                          `CENEIGHCONF` == 3 ~ "Trust most of the People",
                          `CENEIGHCONF` == 6 ~ "Trust all of the People",
                          `CENEIGHCONF` == 96 ~ "Refused",
                          `CENEIGHCONF` == 97 ~ "Don't know",
                          `CENEIGHCONF` == 98 ~ "No Response",
                          `CENEIGHCONF` == 99 ~ "NIU",
                          TRUE ~ "unknown"))
```


Voulgaris edits

```{r}
data_subset <- data %>%
  filter(CENEIGHCONF1 != "NIU") %>%
  mutate(conf_2level = case_when(CENEIGHCONF1 == "Trust most of the People" ~ TRUE,
                                 CENEIGHCONF1 == "Trust all of the People" ~ TRUE,
                                 TRUE ~ FALSE))

svy_data <- data_subset %>%
  as_survey_design(weights = WTFINL)


county_level <- svy_data %>%
  group_by(COUNTY) %>%
  srvyr::summarize(mean = survey_mean(conf_2level),
                   sample = n()) 
```

Get census data

```{r}
county_data <- get_acs(geography = "county", variables = c("B25003_003", "B25003_001"), output = 'wide') %>%
  mutate(pct_rent = B25003_003E / B25003_001E)

### FIX 4-digit FIPS codes by putting the leading zeros back on them,
### the resulting FIPS code needs to be a string (as.character())
### Name the resulting FIPS code GEOID

### remand couty

all_data <- left_join(county_level, county_data)

```
```{r}
paste0("0", county_level$COUNTY)
```

```{r}
county_level1 <- county_level %>% 
  mutate(COUNTY = paste0("0", county_level$COUNTY))
```
